---
title: "Orchard Project: metadata"
output: html_document
---

#### About this document

This is an [R Markdown document](http://rmarkdown.rstudio.com).

## Format metadata for analysis

This project takes place over the course of two days, so changes can be characterized
as part of the constantly increasing vector of time, or as a function of the diurnal
cycle. Environmental factors, like sunlight and rain, also very over the arrow
and cycle of time.

Information about time and environment are in separate file. The goal of this
document is to import and parse these into a usable format.

### Setup:
```{r setup, include=F}
library("checkpoint")
library("knitr")
checkpoint("2016-10-01", use.knitr = T)

library("ggplot2")
library("readr")
library("dplyr")
library("readxl")
library("tibble")
library("lubridate")
theme_set(theme_bw())
opts_chunk$set(cache=F)

```

### Import files


```{r importdata}
meta.fromJim <- read_excel("./orchard_jjm5Oct2016.xlsx", sheet = "sampleInfo")
meta.fromJim

meta.weather <- read_excel("./Weather Data Summary2012_2014.xlsx",
                           sheet = "2014 hourly_2-1 through 10-31")
meta.weather

```



### Improve weather metadata

Let's construct an useful and unambiguous date-time from the weather metadata file.

The column will have the format `yyyy-mm-dd hh:mm:ss UTC`, however only `yyyy-mm-dd hh`
are accurate and useful.

Note that I've choosing to round the minutes to the nearest hour (round `18:41`
to `19:00`), so that we can elegantly match these to measurements taken by our team.

```{r fix-weaather}

head(meta.weather$Date)
# The $Date column is correct enough

head(meta.weather$Time)
# From the $Time column, only the minutes and seconds are correct. So just use those.

summary(minute(meta.weather$Time)) # Measurements take on the 41st minute...
head(hour(meta.weather$Time)) #... starting at midnight and going from there.


# Make a new column combining the correct parts of $Date and $Time
# and name it collection_timestamp to match the other metadata
meta.weather$collection_timestamp <- meta.weather$Time
year(meta.weather$collection_timestamp) <- 2014
month(meta.weather$collection_timestamp) <- month(meta.weather$Date)
day(meta.weather$collection_timestamp) <- day(meta.weather$Date)

# Round (up in this case) to nearest hour
meta.weather$collection_timestamp <- round_date(meta.weather$collection_timestamp, unit = "hour")

# Remove blank columns
meta.weather
meta.weather$`Julian Day` <- NULL
meta.weather$`TMP ° F (Daily Ave)` <- NULL
meta.weather$`Lake Stage` <- NULL
meta.weather$`Lake vol` <- NULL
meta.weather$`Surface area` <- NULL
meta.weather$`Surface area` <- NULL
meta.weather$`Wetted Area` <- NULL
meta.weather

tail(table(meta.weather$collection_timestamp, useNA = "always"))
# Lot's of NAs remain in this data set. Let's drop them.
meta.weather <- na.exclude(meta.weather)

# Now that we have used the accurate part of the Time column, let's drop it too
meta.weather$Time <- NULL

# Also add a column with TMP in C
#meta.weather$`TMP (C)` <- (meta.weather$`TMP ° F` - 32) * (5/9)
meta.weather$`TMP (C)` <- round((meta.weather$`TMP ° F` - 32) * (5/9), digits = 1)
unique(sort(meta.weather$`TMP (C)`))

# Remove unneeded `TMP ° F` column
meta.weather$`TMP ° F` <- NULL
meta.weather

```

### Improve metadata from Jim

Fix import from excel.

```{r fix-jim}
meta.fromJim

# This column could contain light reads from the field site itself, but I have
# not imported this data from Hans.
meta.fromJim$`incident irradaince` <- NULL

# Remove empty Temperature column
meta.fromJim$`Temperature (ᵒC)` <- NULL

```

### Combine metadata

Add matching weather data to data provided by Jim.

Also get hourly rate of organic labeled carbon change using `diff_Ave_d13C-organic`.

```{r combine-meta}
# We will found collection_timestamp so they match up.
meta.fromJim$collection_timestamp <- round_date(meta.fromJim$collection_timestamp, unit = "minute")

meta.weather
head(meta.weather$collection_timestamp)

meta.join <- left_join(x = meta.fromJim, y = meta.weather, by = "collection_timestamp")
meta.join
```

### Improve metadata

Lets convert `diff_Ave_d13C-organic` into a per-hour measurement. (`fin_diff_Ave_d13C-organic` is already per-hour)

Also, split `fin_diff_Ave_d13C.organic` into new vectors per `Substrate`.

```{r improve-meta}
summary((meta.join))
summary(meta.join$`diff_Ave_d13C-organic`)
head(meta.join$`diff_Ave_d13C-organic`)

# divide by hours and round
meta.join$`diff_Ave_d13C-organic` <- meta.join$`diff_Ave_d13C-organic` / 6
# Also divide diff_Stdev by hours
meta.join$diff_Stdev <- meta.join$diff_Stdev / 6


head(meta.join$`diff_Ave_d13C-organic`)
head(meta.join$`fin_diff_Ave_d13C-organic`)
# Much more similar. Good.

# Split Substrate + fin_diff_Ave_d13C.organic into new vectors
meta.join$pp <- meta.join$`fin_diff_Ave_d13C-organic`
meta.join$hpa <- meta.join$`fin_diff_Ave_d13C-organic`
meta.join$hpg <- meta.join$`fin_diff_Ave_d13C-organic`

# For our calculated catories of pp, hpa, and hpg what should we set for the
# samples outside of this group? We use 0 (zero) here instead of NA as it
# works better with the vegan::capscale() function and does not largely change
# the output when passed with taxa vectors.
meta.join$pp[which(meta.join$Substrate != "13C DIC")] <- 0
meta.join$hpg[which(meta.join$Substrate != "13C Glucose")] <- 0
meta.join$hpa[which(meta.join$Substrate != "13C Acetate")] <- 0

# Finally, inspect the Error propagation
head(meta.join$diff_Stdev)
head(meta.join$`fin_diff_Stdev_d13C-organic`)

names(meta.join)
```

## Export

Save it as flat text file for use in downstream analysis

```{r export}

write_tsv(meta.join, "generated-meta.tsv", col_names = T)

```
